### 2024-04-30 paper.tex
update contributions, fix references

\usepackage[utf8]{inputenc}
\usepackage[super,sort&compress]{natbib}
\bibliographystyle{simon}
\setlength{\bibsep}{0.3ex plus 0.2ex}
%\setcitestyle{aysep={}} 
\usepackage[dvipsnames]{xcolor}
\usepackage[compact]{titlesec}
\usepackage[hypertexnames=false]{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={blue!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!50!black}
}
\urlstyle{same} % no tt font for URLs
\renewcommand{\topfraction}{0.9}
\renewcommand{\dbltopfraction}{0.9}
\renewcommand{\textfraction}{0.1}
\clubpenalty=1000
\widowpenalty=1000
\displaywidowpenalty=1000

@@

\section*{Introduction}

@@

Fig.~\ref{fig:dmr}D-E depicts an exemplary DMR, located at the gene body of myelin-basic protein (\textit{Mbp}), the major component of myelin that is essential for myelination of neuronal axons \citep{mbp}.
Our results suggest that oligodendrocyte-specific gene expression of \textit{Mbp} is supported by low methylation at the detected DMR.

@@

\subsection*{Acknowledgments}
SA and LPMK acknowledge funding by the Klaus Tschira Foundation (project 00.022.2019).
LPMK, MMB, LK, SC, and AMV acknowledge funding from the European Commission via ERC grant 771376, from the DFG via SFB grant 873, and from the DKFZ.
We thank Alexey Uvarovskii for code refactoring and helpful suggestions on the source code.

@@

MMB developed and implemented DMR detection under supervision of LPMK and SA.
SO, LK and LPMK performed benchmarks.
LPMK and SA wrote the manuscript.
% TO DO: complete this

@@

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=.8\textwidth]{figures/Fig_DMRs.png}
	\end{center}
	\caption{\small \textbf{Detection of differentially methylated regions (DMRs).}\\
		\textbf{(A)} To identify DMRs, the chromosome is first divided into windows overlapping by a small and fixed size.
		For each window, the cell’s methylation values are calculated as described and as depicted in Fig.~\ref{fig:smoothres}B.
		Then the t statistic of these values is obtained as a measure of methylation difference between two groups of cells that were selected by the user (each point represents one of the windows).
		Upper and lower thresholds (dashed lines) are chosen such that a defined quantile of windows have a t statistic above or below the threshold.
		Windows exceeding either threshold are merged if they overlap, yielding DMRs that are lowly methylated in either group of cells.
		\textbf{(B)} DMRs detected between 58 oligodendrocytes (ODs) and 130 neural stem cells (NSCs) from \citet{kremer_scnmt} (top) and DMRs detected in the same data with randomly permuted cell labels (bottom, used to estimate FDR and determine adjusted p-values).
		\textbf{(C)} Enrichment of GO terms associated with DMRs lowly methylated in ODs (top) or NSCs (bottom).
		Depicted are the top 5 GO terms of the "Biological Process" (GO:BP) and "Cellular Component" (GO:CC) GO category, and their binomial test p-value and enrichment, as reported by GREAT \citep{mclean2010great}.
		\textbf{(D)} Mean methylation of NSCs and ODs at an exemplary DMR.
		Each point corresponds to a cell, black lines denote the median.
		\textbf{(E)} Detailed view of the DMR (yellow dashed rectangle) from C in pseudo-bulk samples consisting of NSCs or ODs.
		Vertical bars represent CpG sites.
	}
	\label{fig:dmr}
\end{figure*}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[super,sort]{natbib}
%\bibliographystyle{simon}
\bibliographystyle{unsrtnat}
\setlength{\bibsep}{0.3ex plus 0.2ex}
%\setcitestyle{aysep={}} 
\usepackage[dvipsnames]{xcolor}
\usepackage[compact]{titlesec}
\usepackage[hypertexnames=false]{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={blue!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!50!black}
}
\urlstyle{same} % no tt font for URLs
\renewcommand{\topfraction}{0.9}
\renewcommand{\dbltopfraction}{0.9}
\renewcommand{\textfraction}{0.1}
\clubpenalty=1000
\widowpenalty=1000
\displaywidowpenalty=1000

@@

\section*{Main Text}
\section*{Introduction}

@@

Fig.~\ref{fig:dmr}D-E depicts an exemplary DMR, located at the gene encoding myelin-basic protein (\textit{Mbp}), the major component of myelin that is essential for myelination of neuronal axons \citep{mbp}.
Our results suggest that oligodendrocyte-specific gene expression of \textit{Mbp} is supported by low methylation at the detected DMR.

@@

\subsection*{Acknowledgements}
SA and LPMK acknowledge funding by the Klaus Tschira Foundation (project 00.022.2019).
LPMK, MMB, LK, SC, and AMV acknowledge funding from the European Commission via ERC grant 771376 (to AMV), from the DFG via SFB 873, and from the DKFZ.
We thank Alexey Uvarovskii for code refactoring and helpful suggestions on the source code.

@@

LPMK and SA conceived and worked out the method.
LPKM wrote the software package, with contributions from MMB (differential methylation functionality) and LKü.
LPMK, SO and LKü prepared the use-case demonstrations and performed the benchmarks.
SC and AMV contributed experimental data and biological expertise.
LPMK, AMV and SA contributed supervision and project management.
LPMK and SA wrote the manuscript.

@@

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=.8\textwidth]{figures/Fig_DMRs.png}
	\end{center}
	\caption{\small \textbf{Detection of differentially methylated regions (DMRs).}\\
		\textbf{(A)} To identify DMRs, the chromosome is first divided into windows overlapping by a small and fixed size.
		For each window, the cell’s methylation values are calculated as described and as depicted in Fig.~\ref{fig:smoothres}B.
		Then the t statistic of these values is obtained as a measure of methylation difference between two groups of cells that were selected by the user (each point represents one of the windows).
		Upper and lower thresholds (dashed lines) are chosen such that a defined quantile of windows have a t statistic above or below the threshold.
		Windows exceeding either threshold are merged if they overlap, yielding DMRs that are lowly methylated in either group of cells.
		\textbf{(B)} DMRs detected between 58 oligodendrocytes (ODs) and 130 neural stem cells (NSCs) from \citet{kremer_scnmt} (top) and DMRs detected in the same data with randomly permuted cell labels (bottom, used to estimate FDR and determine adjusted p-values).
		\textbf{(C)} Enrichment of GO terms associated with DMRs lowly methylated in ODs (top) or NSCs (bottom).
		Depicted are the top 5 GO terms of the "Biological Process" (GO:BP) and "Cellular Component" (GO:CC) GO category, and their binomial test p-value (two-sided, not adjusted for multiple comparisons) and enrichment, as reported by GREAT \citep{mclean2010great}.
		\textbf{(D)} Mean methylation of NSCs and ODs at an exemplary DMR.
		Each point corresponds to a cell, black lines denote the median.
		\textbf{(E)} Detailed view of the DMR (yellow dashed rectangle) from C in pseudo-bulk samples consisting of NSCs or ODs.
		Vertical bars represent CpG sites.
	}
	\label{fig:dmr}
\end{figure*}
